---
title: "Assignment coursera"
author: "bellemin ronan"
date: "10/17/2021"
output: 
  html_document:
    keep_md: true
editor_options: 
  chunk_output_type: inline
---

# Assignment

## Loading and preprocessing the data

Loading packages
```{r, echo=TRUE}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(cowplot)
```

Loading the dataset and dealing with NAs
```{r, echo=TRUE}
dataset <- read_csv("activity.csv")
dataset <- drop_na(dataset)
```

## What is mean total number of steps taken per day?

Calculating the mean of steps per day
```{r, echo=TRUE}
print(mean_day <- dataset %>% 
  group_by(date) %>% 
  summarise(mean_steps = mean(steps)))
```

Histogram of the total number of steps taken each day
```{r, echo=TRUE}
cum_steps_df <- dataset %>% 
  group_by(date) %>% 
  summarise(cum_steps = cumsum(steps))

total_steps_per_day_df <- cum_steps_df %>% 
  group_by(date) %>% 
  summarise(total_steps = max(cum_steps))

ggplot(data = total_steps_per_day_df) +
  geom_histogram(aes(total_steps), col = "black", fill = "brown")
```

Mean and median of the total number of steps per day
```{r, echo=TRUE}
print(mean(total_steps_per_day_df$total_steps, na.rm = TRUE))

print(median(total_steps_per_day_df$total_steps, na.rm = TRUE))
```

## What is the average daily activity pattern?

Time serie plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r, echo = TRUE}
unique_interval <- dataset %>% 
  group_by(interval) %>% 
  summarise(mean_steps = mean(steps))

ggplot(unique_interval)+
  geom_line(aes(interval, mean_steps), col = "blue")
```

To see which 5-min interval contains the most averaged steps, I'm going to look in the table and look around the 800th interval the one that is more than 200 steps away.
It is the 835th 5-min interval.

## Imputing missing values

Reloading original dataset 
```{r, echo=TRUE}
dataset <- read_csv("activity.csv")
```

The total number of Nas in the original dataset
```{r, echo=TRUE}
sum(is.na(dataset))
```

Computing NAs as the mean for that 5-minute interval
```{r, echo=TRUE}
dataset <- dataset %>% 
  left_join(unique_interval, by = "interval")

dataset$steps <- ifelse(is.na(dataset$steps), dataset$mean_steps, dataset$steps)
```

Creating a new dataset same as the original but with NAs filled in
```{r, echo=TRUE}
new_dataset <- dataset
```

Histogram + mean + median total number of steps taken per day
```{r, echo=TRUE}
cum_steps_df_2 <- new_dataset %>% 
  group_by(date) %>% 
  summarise(cum_steps = cumsum(steps))

total_steps_per_day_df_2 <- cum_steps_df_2 %>% 
  group_by(date) %>% 
  summarise(total_steps = max(cum_steps))

ggplot(data = total_steps_per_day_df_2) +
  geom_histogram(aes(total_steps), col = "black", fill = "brown")
```

Calculating the mean and median
```{r, echo=TRUE}
print(mean(total_steps_per_day_df_2$total_steps, na.rm = TRUE))

print(median(total_steps_per_day_df_2$total_steps, na.rm = TRUE))
```

So it seems that imputing NAs as the mean for that 5-minute interval does not have a large impact on the estimates. Indeed, the estimates are still the same, except for the median which has a little bit increased after imputing data.

## Are there differences in activity patterns between weekdays and weekends?

Creating a new factor variable with either weekday or weekend
```{r, echo=TRUE}
print(new_dataset<- new_dataset %>% 
  mutate(weekday = case_when(
    weekdays(date) %in% c("Lundi", "Mardi", "Mercredi", "Jeudi", "Vendredi") ~ "weekday",
    weekdays(date) %in% c("Samedi", "Dimanche") ~ "weekend"
  )))
```
Making a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)
```{r, echo=TRUE}
#seperate in 2 dataframes: one with only weekend and aonther one with weekdays
only_weekend <- filter(new_dataset, weekday == "weekend")
only_weekday <- filter(new_dataset, weekday == "weekday")

#for weekdays
panel_plot_df_1 <- only_weekday %>% 
  group_by(interval) %>% 
  summarise(mean_steps_weekday = mean(steps))

#for weekends
panel_plot_df_2 <- only_weekend %>% 
  group_by(interval) %>% 
  summarise(mean_steps_weekend = mean(steps))

#combine dataframes
final_df<- cbind(panel_plot_df_1, panel_plot_df_2)
final_df <- final_df %>% 
  select(-3)

#create plot
weekday_plot<- ggplot(final_df)+
  geom_line(aes(interval, mean_steps_weekday), col = "blue")

weekend_plot <- ggplot(final_df)+
  geom_line(aes(interval, mean_steps_weekend), col = "blue")

plot_grid(weekday_plot, weekend_plot)
```


